---
title: "FrugBirds"
author: "Iago Ferreiro & Iciar Jiménez"
date: "2022-09-23"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data cleaning AVONET database


```{r load libraries}
library(dplyr)
library(ggplot2)
library(stringr)
```



Import AVONET  BirdLife data 
```{r}
AVONET <- read.csv("Data/Traits/AVONET/AVONET_BirdLife_Data.csv", sep=";")

```


Import crosswalk between BirdLife dataset and BirdTree (Jetz phylo)
```{r}
crosswalk <- read.csv("Data/Traits/AVONET/BirdLife_BirdTree_crosswalk.csv", sep=";")
crosswalk <- crosswalk %>% dplyr::select(BirdLife_Species, BirdTree_Species)

```


Merge with AVONET dataset
```{r}
AVONET <- AVONET %>% dplyr::rename(BirdLife_Species=Species1) 
AVONET <- left_join(AVONET, crosswalk, by="BirdLife_Species")
rm(crosswalk) #clean environment
```

Select only traits of interest
```{r}
AVONET <- AVONET %>% dplyr::select(Avibase.ID1, Family1, Order1, BirdLife_Species,
                            BirdTree_Species, Beak.Length_Culmen,
                            Beak.Length_Nares, Beak.Width, Beak.Depth,
                            Kipps.Distance, Primary.Lifestyle, Centroid.Latitude,
                            Centroid.Longitude, Range.Size) %>%
                   dplyr::rename(ID_Avibase = Avibase.ID1, Family=Family1, Order= Order1,
                           Beak_CulmenLength = Beak.Length_Culmen,
                           Beak_NaresLength = Beak.Length_Nares,
                           Beak_Width = Beak.Width, Beak_Depth= Beak.Depth,
                           KippsD = Kipps.Distance, Lifestyle = Primary.Lifestyle,
                           Centre_Lat = Centroid.Latitude, Centre_Long=Centroid.Longitude,
                           Range_Size = Range.Size)
```

Create genus column

```{r}
AVONET$Genus <- word(AVONET$BirdTree_Species, 1)
```

# Data cleaning EltonTraits database

Load specific libraries
```{r}
library(dplyr)

```


Import Elton Traits database
```{r}
Elton_raw <- read.delim("Data/Traits/ELTON/Elton_BirdFuncDat.txt")

```


Our main aim is to predict those values related to the percentage of fruits
within species-specific diet that were previously interpolated by EltonTraits
owners and curators.
```{r}
table(Elton_raw$Diet.Certainty) 
```
Only Category A is valid for our purposes


Create 2 subsets:
```{r}
Elton <- Elton_raw %>% dplyr::filter(Diet.Certainty=="A") %>% 
                       dplyr::select(Scientific, Diet.Fruit) %>%
                       dplyr::rename(Species = Scientific,
                              Frugivory = Diet.Fruit)


Elton_pred <- Elton_raw %>% dplyr::filter(Diet.Certainty!="A") %>% 
                         dplyr::select(Scientific, Diet.Fruit) %>%
                       dplyr::rename(Species = Scientific,
                       Frugivory_Interpolated = Diet.Fruit) 

```


two last rows from Elton_pred datasets are empty rows
```{r}
length(Elton_pred$Species)
Elton_pred <- Elton_pred[1:2060,]

rm(Elton_raw) #Clean environment

```


Check differences between species names in the two databases
```{r, results = FALSE}
setdiff(Elton$Species, AVONET$BirdLife_Species)  #1491 species names are not coincident

setdiff(Elton$Species, AVONET$BirdTree_Species)  #4 species names are not coincident

setdiff(Elton$Species, AVONET$BirdTree_Species) %in% AVONET$BirdLife_Species  

```


Campylopterus curvipennis and C. excellens are now Pampa curvipennis and P.excellens
```{r}

Elton[which(Elton$Species=="Campylopterus curvipennis"), "Species"]<-"Pampa curvipennis"
Elton[which(Elton$Species=="Campylopterus excellens"), "Species"]<-"Pampa excellens"

```

The other 2 sps are extinct so we delete them in Elton database
```{r}

Elton<- Elton %>% dplyr::filter(Species!= "Philydor novaesi") %>% filter(Species!="Melamprosops phaeosoma")

```

Check again the differences
```{r}
setdiff(Elton$Species, AVONET$BirdTree_Species) #no difference
Elton<-Elton %>% dplyr::rename(BirdTree_Species=Species)
```

Merge both datasets
```{r}
Frug_db <- left_join(Elton, AVONET, by="BirdTree_Species")

```

Duplicated rows appears because there is some uncongruences between BirdTree and
BirdLife species names
```{r, results=FALSE}
duplicated(Frug_db$BirdTree_Species)
Frug_db<- Frug_db[-which(duplicated(Frug_db$BirdTree_Species)),]

```


Add a binomial variable depending if fruits are the major component of the diet (>50%) or not
```{r}
Frug_db<- Frug_db %>% mutate (Frugivore=ifelse(Frugivory>=50,"Yes", "No"))
```


check sps names in prediction database
```{r, results=FALSE}
setdiff(Elton_pred$Species, AVONET$BirdLife_Species)

setdiff(Elton_pred$Species, AVONET$BirdTree_Species) #5 sps names are not coincident

setdiff(Elton_pred$Species, AVONET$BirdTree_Species) %in% AVONET$BirdLife_Species 
```




Anthus longicaudatus = Anthus vaalensis
Phyllastrephus leucolepis = Phyllastrephus icterinus
Lophura hatinhensis = Lophura edwardsi
Polioptila clementsi = Polioptila guianensis
```{r}
Elton_pred[which(Elton_pred$Species=="Phyllastrephus leucolepis"), "Species"]<-"Phyllastrephus icterinus"
Elton_pred[which(Elton_pred$Species=="Anthus longicaudatus"), "Species"]<-"Anthus vaalensis"
Elton_pred[which(Elton_pred$Species=="Polioptila clementsi"), "Species"]<-"Polioptila guianensis"
Elton_pred[which(Elton_pred$Species=="Lophura hatinhensis"), "Species"]<-"Lophura edwardsi"

```


Hypositta perdita = invalid taxon
```{r}
Elton_pred<- Elton_pred %>% dplyr::filter(Species!= "Hypositta perdita") 

```


Check sps names again
```{r}
setdiff(Elton_pred$Species, AVONET$BirdTree_Species)

```


Prediction database with traits
```{r}
Elton_pred<-Elton_pred %>% dplyr::rename(BirdTree_Species=Species)

Elton_pred<- left_join(Elton_pred, AVONET, by="BirdTree_Species")

```


Check duplicates
```{r, results=FALSE}
Elton_pred[which(duplicated(Elton_pred$BirdTree_Species)),]

Elton_pred <- Elton_pred[-which(duplicated(Elton_pred$BirdTree_Species)),]

```


Clean environment
```{r}
rm(Elton) 
```

# Interactive map of fruit importance in bird's diet worldwide


Load specific packages
```{r}
library(sp)
library(mapview)

```


This script aims at plotting the centroids of distribution rage for each
species worldwide. Size of each coordinate will depict the importance
of fruits in bird's diet in order to test if there is an latitudinal
pattern in the consumption of fruits by birds

Check NA's within data and remove it from data (just for plotting)
```{r}
which(is.na(Frug_db$Centre_Lat))
Frug_Spatial <- Frug_db[-which(is.na(Frug_db$Centre_Lat)),]

which(is.na(Frug_Spatial$Centre_Long))
Frug_Spatial <- Frug_Spatial[-which(is.na(Frug_Spatial$Centre_Long)),]

coords <- cbind(Frug_Spatial$Centre_Long, Frug_Spatial$Centre_Lat)


```

Turn df into SpatialPoinstDataframe
```{r}
Frug_Spatial <- SpatialPointsDataFrame(coords=coords, data = Frug_Spatial,
                                       proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

```



Create interactive map
```{r}
map<-mapview(Frug_Spatial,cex="Frugivory", zcol="Family", legend=FALSE, alpha=2)
map
```


Save map as interactive html
```{r}
mapshot(map, url="Figures/Frugivory_Map.html")
rm(list=c("coords", "Frug_Spatial", "map"))#clean environment

```

# Construct a model with poisson distribution and check it
```{r}
library(lme4)
m5<- glmer(Frugivory~scale(Beak_Width, center = T) + scale(KippsD, center=T) + scale(Centre_Lat,center=T) + (1|Genus), data=Frug_db, family="poisson")

summary(m5)
res_sim<-DHARMa::simulateResiduals(m5)
DHARMa::plotSimulatedResiduals(res_sim)
performance::check_model(m5)
```

